---
title: "ICA"
author: "Andrew Wu"
date: "November 2024"
output:
  html_document: default
  pdf_document: default
extra_dependencies:
- amsmath
---

```{r}
set.seed(42)
df <- read.table("emissionssw.dat", header=TRUE)
head(df)
```
```{r}
m1 <- lm(nox ~ ., data=df)
summary(m1)
plot(m1)
```

```{r}
hist(df$nox)
hist(log(df$nox))
```

```{r}
boxplot(df)
```
Quite a few outliers across all time series. Consider quantile clipping to prevent influence of these observations?
Maybe even remove from data and don't bother trying to predict 
```{r}
plot.ts(df)
lapply(names(df), function(col) {
  acf(df[[col]], main = paste("ACF for", col))
})

plot(df$nox[1:6])
acf(log(df$nox))
par(mfrow=c(2,1))
plot(df$noxem[1:13], type='l')
plot(df$nox[1:13], type='l')
```
```{r}
summary(df)
```
lag 1/5 have the highest autocorrelations, with the rest being lower. Makes sense given martingale property + same time of day at lag-5 mark. Since lag 1 has such high autocorrelation, may be worth trying a model fitted on first order differences.


```{r}
pairs(df)
```

```{r}
m1 <- lm(nox ~ ., data=df)
summary(m1)
plot(m1)
# 
# m1 <- lm(nox ~ ., data=df, subset=-c(156, 160, 161))
# summary(m1)
# plot(m1)
```
looks like a quadratic u shaped relationship between residuals and fitted. Let's try squaring humidity since that was insignificant and looked nonlinear
```{r}
m2 <- lm(nox ~. + I(humidity^2), data=df)
summary(m2)
plot(m2)
```
Didn't help, what about windspeed since that was nonlinear too. Still seeing a larger fitted value gives larger variance. Polynomials degree two don't seem to work, though there is some improvement. Common theme is that humidity doesn't seem to be important.
```{r}
# try quantile clipping on data to deal with outliers. Park idea for now
# library(raster)
# dfclipped = clip.data(select(df, c('noxem', 'ws')), lower=0.01, upper=0.99)
# print(dfclipped)
# 
# logging the response gives a good residuals vs fitted plot
# Attempt to fit WLS
# add convergence criterion
max_iter = 100
m3w <- lm(log(nox) ~ (noxem + ws)^2 + I(ws^2) + I(noxem^2), data=df)
for (i in 1:max_iter) {
  m3w <- lm(log(nox) ~ (noxem + ws)^2 + I(ws^2) + I(noxem^2), data=df, weights = 1/abs(resid(m3w)))
}
summary(m3w)
plot(m3w)
```

```{r}
library(MASS)
mlog <- lm(log(nox) ~ (noxem + ws)^2 + I(ws^2) + I(noxem^2), data=df)
summary(mlog)
plot(mlog)
mlog_robust <- rlm(log(nox) ~ (noxem + ws)^2 + I(ws^2) + I(noxem^2), data=df)
summary(mlog_robust)
plot(mlog_robust)
```

More insights from question:
- measurements sequential over one year, sorted by day,time. Need to take this dependence into account
- hypothesis: noxem/ws are important, noxem is the source, ws is the dispersion. Humidity may have an impact, but so far it's not showing.

General:
- train/test set, CV? for model selection/evaluating
performance while checking for overfitting

Dealing with time dependence:
- adding lagged/smoothed/rolling mean versions of variables (can try lag 1 given high autocorr)
- adding time index/seasonal predictor? (low priority)

Tried first order differencing, not really any that much better than the 
striaght linear model it seems. Still the same U-shaped residual problem.
```{r}
df["dy"] <- c(NaN, diff(df$nox))
# m4 <- lm(dy ~ noxem + ws + humidity, data=df)
m4 <- lm(dy ~ (noxem + ws)^2 + I(ws^2) + I(noxem^2), data=df)
summary(m4)
plot(m4)
```

```{r}
m <- lm(log(nox) ~ log(noxem) + ws,data=df)
summary(m)
plot(m)
```

```{r}
m <- lm(log(nox) ~ log(noxem) + ws + ws:log(noxem) + cos(seq(0, 2*pi, length.out=2022)),data=df)
summary(m)
plot(m)
```

```{r}
library(robustbase)
m <- lmrob(log(nox) ~ log(noxem) + ws, data=df)
summary(m)
plot(m)
```